Handling String-Number Conversion in String Constraint Solving

### 2019-11-20 draft.tex
some minor changes

But if array indices are modelled as strings, we must handle arithmetic somehow. Let us look at the case of line 6, we need to update the value of \texttt{x["03"-1]}. The evaluation of the expression \texttt{"03"-1} involves an implicit type conversion from the string \texttt{"03"} to an integer value $3$ due to the \texttt{-} (minus) operation. The result of the evaluation of \texttt{"03"-1} is the integer $2$, which is then converted back to string \texttt{"2"} and used as the array index. Hence \texttt{x["03"-1]} means the array element of \texttt{x["2"]}. Even for a simple example like this, the conversion between string and number is unavoidable. This is a rather basic array operation in JavaScript, and so having only limited capability for handling string-number conversion operations will cripple any analysis of non-trivial JavaScript code.

@@

to explain the main ideas behind our decision procedure.  To make our terminology explicit: $\Phi$ states that $x$ concatenated with $y$ is the same as $y$ with $x$, $n$ is the numeric value of the string $x$, $n > 3$, $y$ is a string of 1's and 2's.  Notice that $\Phi$ is satisfiable. E.g., it has a model $x=``12''$ and $y=``1212''$.

@@

An \emph{integer constraint} $\phi_i$ is a linear constraint over the variables in $\varn$ and length of a word $|x|$ for all $x \in \vars$.
%Formally, assume that $\vars=\{x_i,\ldots,x_n\}$ and $\varn=\{y_i,\ldots,y_m\}$, a length constraint is of the form $(\sum_{i\in[1,n]}(j_i \times \sti{x_i}+k_i\times |x_i|)+ \sum_{i\in[1,m]}(l_i \times y_i)) \odot k$, where $\odot \in \{>,\geq, =, \leq, <\}$, and $j_i,k_i,l_i,k\in \mathbb{Z}$. We define $\modelsof{\phi}= \{I \mid (\sum_{i\in[1,n]}(j_i \times \sti{I(x_i)}+k_i\times |I(x_i)|)+ \sum_{i\in[1,m]}(l_i \times I(y_i))) \odot k \}$.
We define $\modelsof{\phi_i}= \{I \mid I(\phi_i)= \true \}$. 
\lh{should we somehow define the sematntics of the lengt constraints more precisely? That is, how is the value of $|x|$ related to the value of $x$. ?}

@@

Last, we define the encoding function as a counterpart of the decoding. Given a PFA $\pa$ and a language $L\subseteq \Sigma^*$, the set of \emph{$\pa$-encodings} of the words in $L$ is defined as
$\encode \pa (L) = \{e \mid \decode \pa (e) \in L\}$. 
%
Notice that only the words from $L\cap \semof \pa$ have an $\pa$-encoding, 
hence $\encode \pa(L) = \encode \pa(L\cap\semof \pa)$, and so due to Corollary~\ref{corollary:pfa}, $\decode \pa({\encode \pa}(L)) = L\cap\semof \pa$.
%$\encode A (L) = \{(I\cup\pim) \mid \decode A (I\cup \pim) \in L \cap \semof{A}\}$. 
%We have that $\decode A \circ \encode A$ equal the identity function.
%

@@

To formalise this correctness argument, we define an additional notation.
For an interpretation $I$ of string variables, 
we define its $R$-encoding as the set $\encode R(I) = \bigcup_{x\in X} \encode \pa(I(x))$ of interpretations of $V_R\cup \#V_R$, containing combinations of interpretations of character variables of each $R(x)$ that encode the $I$.
%
Further, if $I \cup I'$ is an ainterpretation of a general string constraint, that is, an interpretation $I$ of string variables $X$ united with an interpretation $I'$ of integer variables $N$,
%
then its $R$-encoding is defined as the set $\encode R {(I\cup I')} \defeq \{e \cup I' \mid e\in \encode R(I)\}$ of interpretations of variavbles $V_R \cup \#V_R \cup N$. 
%
We lift this notation to sets of interpretation in the standard manner, i.e., $\encode R (J) = \bigcup_{I \in J}\encode R (I)$ for a set of interpretations $J$.
%
The inductive argument needed in the proof of the correctness of the under-approximation then reads as
$$\semof{\underf R \phi}_{V_R\cup V_{\#R}}  = \encode R{\semof{\phi}}$$ 
%
In the next sections, 
we will formulate the corresponding lemma for linear formulas constructed from each type of atomic string constraints.
%
We note that the restriction of the semantics of $\underf R \phi$ here is needed since $\underf R \phi$ is constructed with some auxiliary variables.

@@

We now describe the construction of $\underf R {\phi_e}$ for an equality constraint $\phi_e\defeq x_0\cdot x_1 \cdots x_n = x_{n+1}\cdot x_{n+2} \cdots x_m$. 
%
To simplify the presentation, we assume that the variables are pairwise different, i.e. that $i\neq j \implies x_i \neq x_j$. We may make this assumption without loss of generality, since multiple occurrences of variables in $\phi_e$ can be eliminated: whenever $x_i = x_j$ for $i \neq j$, we may replace $x_j$ by a fresh variable $x_j'$ and conjoin the modified $\phi_e$ with a new equality $x_j = x_j'$. 
%
We also assume that all disequality constraint $t \neq t'$ are already converted to equivalent equality constraints and integer constraints in the standard way~\cite{abdulla2015norn}.
%

@@

For every length variable $|x|\in X$, 
we add a set of fresh auxiliary variables $\{|v|\mid v\in V_x\}$ to express the length by which $v$ contributes to the length of an $R$-encoded word. 
That is, $|v|$ will be $0$ if $v$ is assigned $\noepsilon$,
otherwise it will equal the number $\#v$ of its occurrences in the word:
$$
\Psi_{\mathsf{|v|}} \defeq (v = \noepsilon \land |v|=0) \lor (v \neq \noepsilon \land |v| = \# v)
$$
The lenght of the encoded word $x$ is then the sum of the lengths contributed by the individual character variables in $V_x$, 
hence we let 
$$
\Psi_{|x|} \defeq |x| = \sum_{v\in V_x} |v| \land  \bigwedge_{v\in V_x} \Psi_{\mathsf{|v|}}\ .
$$
Finally, the linear formula created for $\phi_l$ is 
$$
\underf R {\phi_{l}} \defeq \phi_l \land \bigwedge_{|x|\in X} \Psi_{|x|}
$$
and the following lemma states its correctness.
\begin{lemma}\label{lemma:eqcorrect}
$\semof {\underf R {\phi_l}}_{V_R  \cup \#V_R} = \encode R (\semof{\phi_l})$
\end{lemma}

@@

It is not difficult to see from the examples above that, if $R(x)$ is an arbitrary PFA with $m$ loops, the formula that defines the number $n$ contains at least $m$ exponential components, one for each loop.
To the best of our knowledge, the satisfiability problem of integer constraints with a mix of polynomials and exponentials is still open. As said in the introduction, the problem is difficult even for the case that variables are real numbers. For example, the algorithm in~\cite{kincaid2019closed} involves a quantifier elimination procedure which is double-exponential to the length of the input formula and hence cannot handle large instances. We therefore do not expect that such constraint can be solved efficiently. Instead, we will define a special form of the flat restriction $R(x)$ of $x$ that leads to an easies linear formula.

\paragraph{PFA that allows efficient string-number conversion: } 
%Now we know that it is unlikely to use arbitrary PFA and gain efficient solving procedure, due to the requirement to solve exponential integer constraints. The cause of the problem is that each loop in the PFA creates an exponential component in the constraint. 
%An immediate idea would be that maybe we can some pick some specific PFA that allow more efficient constraint solving. 
We will now discuss the special form of $R(x)$, called numeric PFA, which we choose for string variables appearing in string-integer constraints that leads to efficient underapproximation.
%
Besides simple induced linear formulae, we still want single $R(x)$ to cover as many numerals as possible.
We want an ``easy'' completeness property, that is, that the space of all numerals can be covered completely by our special numeric PFAs, and that these numeric PFAs are generated easily, and each of them covers a large and practically significant portion of numerals (so that satifiable assignments can be often found within the domain restriction of just few of numeric PFAs).
%
Particularly, we will proceed towards a definition of \emph{number PFA} $(A^m,\paf^m),m\in\nat$ which covers all numerals that encode integers with $m$ digits.

@@

However, the formula is satisfiable when $x=``00010"$. Observe that $\sti{``00010"}=10$. The key is that even for a bounded integer, the corresponding numeral can be of an unbounded length with arbitrarily many trailing `$0$'s at the front. All numbers with up to $m$-digits can be however still handled without having to solve exponential constraints. It is enough to equip the initial state of the PFA with a  $0$-self-loop.

Consequently, the automaton $A^m$ of our number PFA will have the following form illustrated in Figure~\ref{fig:sfa_its}. 
It has a self-loop on the initial state labeled by the character variable $v_0$, 
forced by the constraint 
$$
\Psi_{v_0}\defeq v_0=0
$$ 
to hold the value $0$. 
This transition ensures that the under-approximation handles numerals with arbitrary number of trailing zeros.  
The self-loop is followed by a chain of $m$ transitions $(q_{i-1},v_i,q_{i}),1\leq i \leq m$, leading towards the final state $q_m$. 
The chain encodes at most $m$ meaningful digits (only \emph{at most} because the first variables in the sequence can still be assigned zeros and some variables may be assigned $\epsilon$). Hence this PFA covers all numerals that encode numbers with at most $m$ digits.
Although it still has a loop, it will not create any exponential component defining the value of $n$ because the loop only represents a sequence of ``0'' at the front of $x$. It will not affect the integer value of $n=\sti{x}$.

@@

%Such PFA handles bounded integers (e.g., 10-digits or 64-bits) precisely and under-approximate the possible values of unbounded integers. It works very efficiently because we just need to use linear constraints to characterize the relation between $x$ and $n$. 
Numeric PFA with these restrictions would satisfy our primary objective, that is, they would induce linear formulae and would ``easily'' and completely cover all numerals.
%
A last problem still needs to be solved beofe they can be efficient in practice. 
%
Recall that the character variables can be assigned $\epsilon$.  
%
Therefore, a single chain of $k$ interesting digits, $k\leq m$, can be by $A^m$ represented in $k\choose m$ ways, each corresponding to one possible interleaving of $k$ digits with $m-k$ epsilons. 
%
\lh{I am not sure whether it is really impossible to make a smlall formula}
%
This would lead to an exponentail expolostion of the size of the linear formula definig the value of $n$.
%
In order to eliminate this blow-up in the size of the formula, we add to $\paf^m$ an additional constraint that forces all epsilons to be \emph{shifted} behind the least significant digit, effectively prohibiting all interleavings but one. This is the formula
$$
\Psi_{\mathsf{shift}}^m \defeq \bigwedge_{1\leq i \leq m} v_i \neq \noepsilon \implies v_{i-1}\neq \noepsilon \ .
$$
Last, since this restriction is meanigfull only when the string is indeed a numeral,  
we also define the constraint representing the strings which are not numerals, the formula
$$
\Psi_{\mathsf{NaN}}^m \defeq  \bigvee_{i\in [1,m]} v_i>9
$$ 
and define the final form of the interpretation restriction used by $A^m$ as 
$$
\paf^m \defeq  \Psi_{\mathsf{NaN}}^m \lor (\Psi_{v_0} \land \Psi_{\mathsf{shift}}^m)\ .
$$
Consequently, we design our domain restrictions $R$ so that for string variables $x$ that appear within string-integer constraints,  $R(x)$ is a number PFA $(A^m,\paf^m),m\in \nat$.

Assuming that the domain restriction for $x$ is $(A^m,\paf^m)$, 
the value of the integer $n$ can be extracted from a numeral using the formula
$$
\Psi_{\mathsf{toInt}}^m\defeq \bigvee_{1\leq k \leq m} \Psi_{\mathsf{last(k)}}^m \land (n = v_1* 10^{k-1} + v_2* 10^{k-2} + \ldots + v_{k})
$$
where $\Psi_{\mathsf{last(k)}}^m$ says that the last variable of $A^m$ assigned a non-$\epsilon$ value is $v_{k}$, namely 
$$
\Psi_{\mathsf{last(k)}}^m \defeq (k = m \land v_k \geq 0) \lor (v_{k} \geq 0 \land v_{k+1} = -1)\ .
$$
Since we also need to distinguish the case when $x$ is not a number, in which case $n$ should equal $-1$, 
the formula underapproximating $\phi_s$ is finally constructed as  
\begin{multline*}
\underf R {\phi_s} 
\defeq 
\parikhfof{A^m} 
\land \\
\bigl((\Psi_{\mathsf{NaN}}^m \land n = -1) 
\lor 
(\neg\Psi_{\mathsf{NaN}}^m \land \Psi_{\mathsf{toInt}}^m)\bigr)
\end{multline*}
The following lemma states correctness of this construction:
\begin{lemma}
$
\semof {\underf R {\psi_s} }_{V_R \cup \#V_R \cup \{n\}} = \encode  R(\semof{\phi_s})
$
\end{lemma}

>>>>

But if array indices are modeled as strings, we must handle arithmetic somehow. Let us look at the case of line 6, we need to update the value of \texttt{x["03"-1]}. The evaluation of the expression \texttt{"03"-1} involves an implicit type conversion from the string \texttt{"03"} to an integer value $3$ due to the \texttt{-} (minus) operation. The result of the evaluation of \texttt{"03"-1} is the integer $2$, which is then converted back to string \texttt{"2"} and used as the array index. Hence \texttt{x["03"-1]} means the array element of \texttt{x["2"]}. Even for a simple example like this, the conversion between string and number is unavoidable. This is a rather basic array operation in JavaScript, and so having only limited capability for handling string-number conversion operations will cripple any analysis of non-trivial JavaScript code.

@@

to explain the main ideas behind our decision procedure.  To make our terminology explicit: $\Phi$ states that $x$ concatenated with $y$ is the same as $y$ with $x$, $n$ is the numeric value of the string $x$, $n > 3$, $y$ is longer than $x$, and $y$ is a string in the form of a finite repetitions of $``12''$.  Notice that $\Phi$ is satisfiable. E.g., it has a model $x=``12''$ and $y=``1212''$.

@@

An \emph{integer constraint} $\phi_i$ is a linear constraint over the variables in $\varn$ and length of a word $|x|$ for all $x \in \vars$.
%Formally, assume that $\vars=\{x_i,\ldots,x_n\}$ and $\varn=\{y_i,\ldots,y_m\}$, a length constraint is of the form $(\sum_{i\in[1,n]}(j_i \times \sti{x_i}+k_i\times |x_i|)+ \sum_{i\in[1,m]}(l_i \times y_i)) \odot k$, where $\odot \in \{>,\geq, =, \leq, <\}$, and $j_i,k_i,l_i,k\in \mathbb{Z}$. We define $\modelsof{\phi}= \{I \mid (\sum_{i\in[1,n]}(j_i \times \sti{I(x_i)}+k_i\times |I(x_i)|)+ \sum_{i\in[1,m]}(l_i \times I(y_i))) \odot k \}$.
We define $\modelsof{\phi_i}= \{I \mid I(\phi_i)= \true \}$.

@@

Last, we define the encoding function as a counterpart of the decoding. Given a PFA $\pa$, and a language $L\subseteq \Sigma^*$, we define the set of $\pa$-encodings of the words in $L$ as 
$\encode \pa (L) = \{e \mid \decode \pa (e) \in L\}$. 
%
Notice that only the words from $L\cap \semof \pa$ have an $\pa$-encoding, 
hence $\encode \pa(L) = \encode \pa(L\cap\semof \pa)$, and so due to Corollary~\ref{corollary:pfa}, $\decode \pa({\encode \pa}(L)) = L\cap\semof \pa$.
%$\encode A (L) = \{(I\cup\pim) \mid \decode A (I\cup \pim) \in L \cap \semof{A}\}$. 
%We have that $\decode A \circ \encode A$ equal the identity function.
%

@@

To this end, we define an additional notation.
For an interpretation $I:X\rightarrow \Sigma^*$, we define its $R$-encoding as the set $\encode R(I) = \bigcup_{x\in X} \encode \pa(I(x))$. We lift this notation to sets of interpretation in the standard manner, i.e., $\encode R (J) = \bigcup_{I \in J}\encode R (I)$ for a set of interpretations $J$.
%
The inductive argument needed in the proof of the correctness of the under-approximation then reads as
$$\semof{\underf R \phi}_{V_R\cup V_{\#R}}  = \encode R{\semof{\phi}}$$ 
%
In the next sections, 
we will formulate the corresponding lemma for linear formulas constructed from each type of atomic string constraints.
%
We note that the restriction of the semantics of $\underf R \phi$ here is needed since $\underf R \phi$ is constructed with some auxiliary variables.

@@

We now describe the construction of $\underf R {\phi_e}$ for an equality constraint $\phi_e\defeq x_0\cdot x_1 \cdots x_n = x_{n+1}\cdot x_{n+2} \cdots x_m$. 
%
To simplify the presentation, we assume that the variables are pairwise different, i.e. that $i\neq j \implies x_i \neq x_j$. We may make this assumption without loss of generality, since multiple occurrences of variables in $\phi_e$ can be eliminated: whenever $x_i = x_j$ for $i \neq j$, we may replace $x_j$ by a fresh variable $x_j'$ and conjoin the modified $\phi_e$ with a new equality $x_j = x_j'$. 
%
We also that all disequality constraint $t \neq t'$ are already converted to equal-satisfiable equality constraints and integer constraints in the standard way~\cite{abdulla2015norn}.
%

@@

For every length variable $|x|\in X$, 
we add a set of fresh auxiliary variables $\{|v|\mid v\in V_x\}$ to express the length by which $v$ contributes to the length of an $R$-encoded word. 
That is, $|v|$ will be $0$ if $v$ is assigned $\noepsilon$,
otherwise it will equal the number $\#v$ of its occurrences in the word:
$$
\Psi_{\mathsf{|v|}} \defeq (v = \noepsilon \land |v|=0) \lor (v \neq \noepsilon \land |v| = \# v)
$$
The lenght of the encoded word $x$ is then the sum of the lengths contributed by the individual character variables in $V_x$, 
hence we let 
$$
\Psi_{|x|} \defeq |x| = \sum_{v\in V_x} |v| \land  \bigwedge_{v\in V_x} \Psi_{\mathsf{|v|}}\ .
$$
Finally, the linear formula created for $\phi_l$ is 
$$
\underf R {\phi_{l}} \defeq \phi_l \land \bigwedge_{|x|\in X} \Psi_{|x|}
$$
\lh{a correctness lemma would be good here, but it needs some modification of encoding probably, to be done}

@@

It is not difficult to see from the examples above that, if $R(x)$ is an arbitrary PFA with $m$ loops, the formula to defines the number $n$ contains at least $m$ exponential components, one for each loop.
To the best of our knowledge, the satisfiability problem of integer constraints with a mix of polynomials and exponentials is still open. As said in the introduction, the problem is difficult even for the case that variables are real numbers. For example, the algorithm in~\cite{kincaid2019closed} involves a quantifier elimination procedure which is double-exponential to the length of the input formula and hence cannot handle large instances. We therefore do not expect that such constraint can be solved efficiently. Instead, we will define a special form of the flat restriction $R(x)$ of $x$ that leads to an easies linear formula.

\paragraph{PFA that allows efficient string-number conversion: } 
%Now we know that it is unlikely to use arbitrary PFA and gain efficient solving procedure, due to the requirement to solve exponential integer constraints. The cause of the problem is that each loop in the PFA creates an exponential component in the constraint. 
%An immediate idea would be that maybe we can some pick some specific PFA that allow more efficient constraint solving. 
We will now discuss the special form of $R(x)$, called numeric PFA, which we choose for string variables appearing in string-integer constraints that leads to efficient underapproximation.
%
Besides simple induced linear formulae, we still want single $R(x)$ to cover as many numerals as possible.
We want an ``easy completeness'' property, that is, that the space of all numerals can be covered completely by our special numeric PFAs, and that these numeric PFAs are generated easily, and each of them covers a large and practically significant portion of numerals (so that satifiable assignments can be often found within the domain restriction of just few of numeric PFAs).
%
Particularly, we will proceed towards a definition of \emph{number PFA} $(A^m,\paf^m),m\in\nat$ which covers all numerals that encode integers with $m$ digits.

@@

However, the formula is satisfiable when $x=``00010"$. Observe that $\sti{``00010"}=10$. The key is that even for a bounded integer, the corresponding string can be of unbounded length with arbitrarily many trailing `$0$'s at the front. All numbers with up to $m$-digits can be however still handled without having to solve exponential constraints. It is enough to equip the initial state of the PFA with a  $0$-self-loop.

Consequently, the automaton $A^m$ of our number PFA will have the following form illustrated in Figure~\ref{fig:sfa_its}. 
It has a self-loop on the initial state labeled by the character variable $v_0$, 
forced by the constraint 
$$
\Psi_{v_0}\defeq v_0=0
$$ 
to hold the value $0$. 
This transition ensures that the under-approximation handles all numerals with arbitrary number of trailing zeros.  
The self-loop is followed by a chain of $m$ transitions $(q_{i-1},v_i,q_{i}),1\leq i \leq m$, leading towards the final state $q_m$. The chain encodes at most $m$ meaningful digits (only \emph{at most} because the first variables in the sequence can still be assigned zeros and some variables may be assigned $\epsilon$). Hence this PFA covers all numerals that encode numbers with at most $m$ digits.
Although it still has a loop, it will not create any exponential component defining the value of $n$ because the loop only represents a sequence of ``0'' at the front of $x$. It will not affect the integer value of $n=\sti{x}$.

@@

%Such PFA handles bounded integers (e.g., 10-digits or 64-bits) precisely and under-approximate the possible values of unbounded integers. It works very efficiently because we just need to use linear constraints to characterize the relation between $x$ and $n$. 
Numeric PFA with these restrictions would satisfy our primary objective, that is, they would induce linear formulae and would easily and completely cover all numerals.
%
A last problem still needs to be solved beofe they can be efficient in practice. 
%
Recall that the character variables can be assigned $\epsilon$.  
%
Therefore, a single chain of $k$ interesting digits, $k\leq m$, can be by $A^m$ represented in $k\choose m$ ways, each corresponding to one possible interleaving of $k$ digits with $m-k$ epsilons. 
%
\lh{I am not sure whether it is really impossible to make a smlall formula}
%
This would lead to an exponentail expolostion of the size of the linear formula definig the value of $n$.
%
In order to eliminate this blow-up in the size of the formula, we add to $\paf^m$ an additional constraint that forces all epsilons to be \emph{shifted} behind the least significant digit, effectively prohibiting all interleavings but one. This is the formula
$$
\Psi_{\mathsf{shift}}^m \defeq \bigwedge_{1\leq i \leq m} v_i \neq \noepsilon \implies v_{i-1}\neq \noepsilon \ .
$$
Last, since this restriction is meanigfull only when the string is indeed numeral,  
we also define the constraint representing the strings which are not numerals, the formula
$$
\Psi_{\mathsf{NaN}}^m \defeq  \bigvee_{i\in [1,m]} v_i>9
$$ 
and define the final form of the interpretation restriction used by $A^m$ as 
$$
\paf^m \defeq  \Psi_{\mathsf{NaN}}^m \lor (\Psi_{v_0} \land \Psi_{\mathsf{shift}}^m)\ .
$$
Consequently, we design our domain restrictions $R$ so that for string variables $x$ that appear within string-integer constraints,  $R(x)$ is a number PFA $(A^m,\paf^m),m\in \nat$.

Assuming that the domain restriction for $x$ is $(A^m,\paf^m)$, 
the value of the integer $n$ can be extracted from a numeral using the formula
$$
\Psi_{\mathsf{toInt}}^m\defeq \bigvee_{1\leq k \leq m} \Psi_{\mathsf{last(k)}}^m \land (n = v_1* 10^{k-1} + v_2* 10^{k-2} + \ldots + v_{k})
$$
where $\Psi_{\mathsf{last(k)}}^m$ says that the last variable of $A^m$ assigned a non-$\epsilon$ value is $v_{k}$, namely 
$$
\Psi_{\mathsf{last(k)}}^m \defeq (k = m \land v_k \geq 0) \lor (v_{k} \geq 0 \land v_{k+1} = -1)\ .
$$
Since we also need to distinguish the case when $x$ is not a number, in which case $n$ should equal $-1$, 
the formula underapproximating $\phi_s$ is finally constructed as  
\begin{multline*}
\underf R {\phi_s} 
\defeq 
\parikhfof{A^m} 
\land \\
\bigl((\Psi_{\mathsf{NaN}}^m \land n = -1) 
\lor 
(\neg\Psi_{\mathsf{NaN}}^m \land \Psi_{\mathsf{toInt}}^m)\bigr)
\end{multline*}
The following lemma states correctness of this construction:
\begin{lemma}
$
\semof {\underf R {\psi_s} }_{V_R \cup \#V_R \cup \{n\}} = \encode  R'(\semof{\phi_s})
$
\end{lemma}
The lemma uses the encoding function generalised to assignments of string \emph{and integer} variables. 
It is defined so that, for an interpretation $I$ of string variables $X$ and an interpetation $I'$ of integer variables $N$, $X\cap N = \emptyset$, 
$\encode R' {(I\cup I')} \defeq \{e \cup I' \mid e\in \encode R(I)\}$, which is then lifted to sets in the standard manner.
